A Dissection of Foregrounds in Epoch of Reionization Observations from the Murchison Widefield Array

### 2015-01-18 ms.tex
Completed section on role of antenna shapes and changed abstract, intro and summary accordingly

Detection of 21~cm emission of neutral hydrogen from the epoch of reionization, at redshifts $z>6$, is limited primarily by foreground emission.  We investigate the chromatic signatures of an all--sky foreground model imprinted by an instrumental transfer function. Using the delay spectrum technique {\bf that maps the Fourier transform of the measurements to locations of foreground objects on the celestial sphere through delays in signals received by antenna pairs}, we demonstrate that the foreground signatures that have the largest impact on the H{\sc i} signal arise from power received far away from the primary field of view. Comparing data from recent Murchison Widefield Array observations with simulations separated into components based on type of emission, we identify diffuse emission near the horizon as a significant contributing factor, even on wide antenna spacings {\bf which are usually not sensitive to diffuse emission}. For signals entering through the primary field of view, compact emission dominates the foreground contamination. These two mechanisms imprint a characteristic {\it pitchfork} signature on the ``foreground wedge'' in Fourier space. Based on these results, we propose that selective down--weighting of data based on antenna spacing vector and time will mitigate foreground contamination substantially (up to a factor $\sim 10$ with negligible loss of sensitivity) in redshifted 21~cm power spectrum analysis.

@@

Considerable effort has been and continues to be made towards understanding the $k$--space behavior of foreground signatures in the observed power spectrum and formulating robust estimators of the true power spectrum \citep{bow09,liu09,dat10,liu11,mor12,tro12,pob13,thy13,dil14,liu14a,liu14b}. {\bf A model that provides a generic explanation for the observed foreground power spectrum has emerged, whereby the wide--field (and chromatic) response of the instrument causes the power in smooth spectrum foregrounds to occupy higher $k$--modes into the so-called `wedge'}. % A general model of scattering of smooth spectrum foregrounds to higher $k$--modes by the wide--field (and chromatic) response of the instrument, into the so-called `wedge', has emerged that provides a high level explanation for the observed foreground power spectrum. 
The conservative foreground strategy that has developed alongside this work is to discard $k$--modes which could be contaminated ({\it avoidance}). The more aggressive alternative is to subtract a sky model and regain access to modes discarded by {\it avoidance}. In both cases, which parts of the sky are most critical to either {\it avoid} or {\it subtract} has remained largely uncertain. Here, we will focus primarily on extending the {\it avoidance} strategy by identifying foreground components at greatest risk to ``leak'' from foreground modes to EoR modes and demonstrating a scheme for down--weighting these components.

@@

After matching real observations from the Murchison Widefield Array with instrumental simulations of the entire known sky, we examine the {\bf shape of the delay spectrum} of each baseline. We report two important findings: the foregrounds that most severely {\bf obscure} the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather by bright objects from near the horizon; and, diffuse Galactic emission plays a significant role hitherto unpredicted. We quantify this contamination by separating the simulations into components based on type of foreground emission. We then arrive at a new method for minimizing the contribution of bright foregrounds which uses prior knowledge of the sky to down--weight adversely contaminated baselines.

In \S\ref{sec:delay-spectrum} we {\bf provide a brief overview of the delay spectrum technique} \citep{par12a,par12b}. In \S\ref{sec:instrument}, we describe the MWA setup, summarize the observing parameters, and present the resulting data. Simulations using these observing parameters are described in \S\ref{sec:modeling} and analyzed by sky location and type of emission in \S\ref{sec:delay-spectrum-analysis}. In \S\ref{sec:fg-grading}, we offer an initial description of a more precise foreground avoidance technique. We present a summary of our work and findings in \S\ref{sec:summary}.

@@

Since we are concerned with a redshifted H{\sc i} spectral line from cosmological distances, $f$ is a measure of cosmological distance along the line of sight. $\eta$, which is the Fourier transform dual of $f$, is used to denote the spatial frequency along the line of sight and has units of time. It is directly related to the line--of--sight wavenumber, 
\begin{align}\label{eqn:k-prll}
  k_\parallel &\approx \frac{2\pi\eta\,f_{21}H_0\,E(z)}{c(1+z)^2}, 
\end{align}
where, $f_{21}$ is the rest frame frequency of the 21~cm spin flip transition of H{\sc i}, and $H_0$ and $E(z)\equiv [\Omega_\textrm{M}(1+z)^3+\Omega_\textrm{k}(1+z)^2+\Omega_\Lambda]^{1/2}$ are standard terms in cosmology. This approximation holds under the assumption that the redshift range (or frequency band) is small enough within which cosmological evolution is negligible. Thus,
\begin{align}\label{eqn:los-transform}
  V_{u\eta} &\equiv V_{u\eta}(\boldsymbol{u},\eta) \nonumber\\
  &\equiv \int V_{uf}(\boldsymbol{u},f)\,W'_f\,e^{i2\pi f\eta}\,\dif f
\end{align}
represents the true spatial Fourier representation of the three--dimensional sky brightness distribution. This approach has been discussed in detail in \citet{mor04}. The spatial power spectrum of EoR H{\sc i} distribution, $P(\boldsymbol{k}_\perp,k_\parallel)$, and $V_{u\eta}$ are related by \citep{mor04,mcq06,par12a}: 
\begin{align}\label{eqn:true-power_spectrum}
  P(\boldsymbol{k}_\perp,k_\parallel) &\simeq |V_{u\eta}|^2\left(\frac{A_\textrm{e}}{\lambda^2\Delta B}\right)\left(\frac{D^2\Delta D}{\Delta B}\right)\left(\frac{\lambda^2}{2k_\textrm{B}}\right)^2,
\end{align}
where, $A_\textrm{e}$ is the effective area of the antenna, $\Delta B$ is the bandwidth, $\lambda$ is the wavelength of the band center, and $k_\textrm{B}$ is the Boltzmann constant. Thus, $V_{u\eta}$ inferred from observations, also in units of Jy~Hz, can be converted into an equivalent cosmological H{\sc i} power spectrum $P(\boldsymbol{k}_\perp,k_\parallel)$, in units of K$^2$~Mpc$^3$.

@@

% Radio emission from foreground objects, due to their spatial and spectral (chromatic) structure, contaminate the received signal \citep{bow09,dat10,mor12,tro12,thy13,liu14a,liu14b}. 
Foregrounds can be described in both the delay spectrum and Fourier frameworks. For our study, which primarily concerns with foreground characterization, we find the former to be simple and yet extremely useful, while maintaining a close correspondence with the latter despite subtle differences. In order to express a quantity derived from $V_{u\eta}$ that has units identical to that of aforementioned cosmological H{\sc i} power spectrum derived, we define:
\begin{align}\label{eqn:fake-power-spectrum}
  P_\textrm{d}(\boldsymbol{k}_\perp,k_\parallel) &\equiv |V_{b\tau}|^2\left(\frac{A_\textrm{e}}{\lambda^2\Delta B}\right)\left(\frac{D^2\Delta D}{\Delta B}\right)\left(\frac{\lambda^2}{2k_\textrm{B}}\right)^2,
\end{align}
where, $\boldsymbol{u}$ and $\eta$ in equations~\ref{eqn:k-perp} and \ref{eqn:k-prll} have been correspondingly replaced with $\boldsymbol{b}/\lambda$ and $\tau$ respectively.

In summary, the delay spectrum, $V_{b\tau}$, is obtained from {\it visibilities} which are the basic data blocks measured by each antenna pair, using equations~\ref{eqn:obsvis} and \ref{eqn:delay-transform}. $V_{b\tau}$ captures all the effects of EoR H{\sc i} signal corruption caused by foregrounds and the instrument. At the same time, it is closely related to the sought power spectrum $P(\boldsymbol{k}_\perp,k_\parallel)$ containing critical information about spatial scales. For purposes of illustration later in this paper, we often use the quantities $|V_{b\tau}|$ and $P_\textrm{d}(\boldsymbol{k}_\perp,k_\parallel)$.

@@

For geometrical intuition, we restrict the orientation ($\theta_b$, measured anti-clockwise from East) of all baselines to lie in the range $-67.5\arcdeg \leq \theta_b < 112.5\arcdeg$. Baselines oriented in the other half--plane measure conjugate visibilities with delays of equal magnitude but of opposite sign and hence are ignored in our analysis.

@@

The {\it local sidereal time} (LST) at these two snapshots are shown in Figure~\ref{fig:pointings} as vertical lines at {\bf 22.08~hours ($-1.92$~hours in the figure)} and 0.09~hours which are hereafter denoted as {\it off--zenith} and {\it zenith} pointings, respectively.

@@

The data are flagged\footnote{\url{http://sourceforge.net/p/aoflagger}} for interference \citep{off10,off12}, removing 3\% of the data and averaged in time  and frequency from the raw 0.5~s, 40~kHz to 2s, 80~kHz. These data are then calibrated to a simulation of the sky containing 2420 point--like objects selected from the MWA Commissioning Survey \citep[MWACS;][]{hur14}. It has a flux density limit of 25~mJy and a declination range of $-12\arcdeg$~ to $-40\arcdeg$~ evenly covering the field of view of the observations reported here. The objects used in calibration are selected to lie inside the 5\% contour of the primary lobe of the tile power--pattern. The calibration algorithm -- based on forward modeling software by \citet{sul12} and the calibration method described by \citet{sal14} -- computes complex gain solutions per channel per antenna averaged to two minute intervals. The solutions are fairly low signal--to--noise so we iteratively average along the antenna and frequency dimensions to capture the relatively independent passband and antenna--to--antenna variation. First, we average the channel gains over all antennas to obtain a high signal--to--noise measurement of the bandpass. After applying this single passband, we do a second round of calibration and fit second and first order polynomials for amplitude and phase respectively for each antenna. This flattens any residual variation in bandpass and removes small phase slopes due to variations in cable delay. Finally, we fit for an additional phase known to be caused by small reflections in subset of cables.

@@

In subsequent sections, we plot the amplitude of the delay spectrum, $|V_{b\tau}|$, sorted by baseline length, $|\boldsymbol{b}|$. Note that visibilities from different baselines have not been averaged together in this analysis.

We show results of delay spectrum from MWA data after deconvolution along the delay axis in Figure~\ref{fig:fhd_data}. The principle feature is the {\it foreground wedge}, whose boundaries, marked by white lines, increase linearly with light travel time corresponding to baseline length. Note that the wedge is not filled out uniformly. Each baseline has a different orientation and resulting response to different objects in the sky. It is this variation that we will explore below.

@@

\begin{figure}[htb]
\centering
% \includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_fhd_delay_spectrum_snapshots.eps}
% \includegraphics[width=\linewidth]{f3.eps}
\includegraphics[width=\linewidth]{fig3.eps}
\caption{Delay spectra power of two observations of the EoR field (RA = 0$^\textrm{h}$, Dec = $-30\arcdeg$) at an {\it off--zenith} pointing (top) when tiles have significant secondary lobes in the direction of the Galactic center and a {\it zenith} pointing (bottom) where secondary lobes are minimal and the Galaxy has set. {\bf In both cases, the delay spectra are phased to the zenith and thus the {\it foreground wedge} boundaries are symmetric around $\tau=0$}. For reference, the $k$--space equivalent of $\tau$ and $|\boldsymbol{b}|$, namely, $k_\parallel$ and $k_\perp$ are shown at the axes on the right and top respectively. White lines mark the boundaries of {\it foreground wedge} determined by light travel time across each baseline. Smooth spectrum objects predominantly reside in this {\it foreground wedge}. The logarithmic color scale (shown at the right) is common to both panels. \label{fig:fhd_data}}
\end{figure}

@@

{\bf Estimation of $T_\textrm{sys}$ from MWA data is a subject of active investigation. For our work, we estimate thermal noise in the data using the {\it rms} of $V_{b\tau}$ obtained after {\it delay--deconvolution} across all antenna spacings for $|\tau| \geq 1$~$\mu$s using the relations:
\begin{align}\label{eqn:Tsys}
  \Delta V_{b\tau}^\textrm{rms} &= \sqrt{N_\textrm{ch}}\,\Delta V_{bf}^\textrm{rms} \Delta f, \textrm{and} \\
  \Delta V_{bf}^\textrm{rms} &= \frac{2\,k_\textrm{B}\,T_\textrm{sys}}{A_\textrm{e}\sqrt{2\,\Delta f\,\Delta t}},
\end{align}
where, $\Delta f = 80$~kHz, $\Delta t = 112$~s, and $N_\textrm{ch}=\Delta B/\Delta f$ is the number of frequency channels. The choice of threshold for $\tau$ is well outside the {\it foreground window}, where foreground contamination is negligible and thus makes a robust estimation of $T_\textrm{sys}$ possible. We find the average system temperature to be $\sim 95$~K across all frequency channels and across all antenna pairs throughout the course of the observation. Hence, for our simulations, we use $T_\textrm{sys}=95$~K to match the thermal noise observed in the data.}

@@

With the aforementioned foreground model, and instrumental and observational parameters, we simulate visibilities using equation~\ref{eqn:obsvis}. Figure~\ref{fig:sim_data} shows the amplitude of the delay spectrum from {\it off--zenith} and {\it zenith} pointings. Notice the qualitative agreement of amplitude and structure with those obtained from data shown in Figure~\ref{fig:fhd_data}. The Galactic center and the Galactic plane visible in the {\it off--zenith} pointing make it appear brighter in the {\it foreground wedge}.

@@

In order to make a quantitative comparison of delay spectra obtained with MWA data and our simulations, we consider the uncertainty in the assumed spectral index of our foreground model. Our foreground models are derived from other higher frequency catalogs and sky maps. The inherent spread in spectral index increases the uncertainty while predicting fluxes at the observing frequency. Using simple error propagation, the fractional error in the delay spectrum {\bf amplitude} caused by the spread in spectral index is $\sim \ln(f_\textrm{orig}/f)\,\Delta\alpha_\textrm{sp}$, where, $f_\textrm{orig}$ is the original frequency at which the catalog or map was created, $f=185$~MHz is the MWA observing frequency, and $\Delta\alpha_\textrm{sp}$ is the spread (HWHM) in spectral index. From \citet{mau03}, we assume $\Delta\alpha_\textrm{sp} \approx 0.35$ for point sources from NVSS and SUMSS catalogs. Although the model of \citet{deo08} yields a spectral index per direction on the sky, we could assume similar uncertainties exist in spectral indices of our diffuse sky model as well, which is predominantly derived from the 408~MHz map of \citet{has82}. Thus, fractional errors in delay spectrum {\bf amplitudes} from compact and diffuse components are $\sim$70\% and $\sim$30\% respectively.

@@

\item {\it Baseline Orientation}: Since the spatial structure of our foreground model is not expected to be isotropic, we divide our antenna spacings by their orientation, $\theta_\textrm{b}$. We use the {\bf four following bins sweeping 45\arcdeg each}: $-67.5\arcdeg\le \theta_\textrm{b} < -22.5\arcdeg$, $-22.5\arcdeg\le \theta_\textrm{b} < 22.5\arcdeg$, $22.5\arcdeg\le \theta_\textrm{b} < 67.5\arcdeg$, and $67.5\arcdeg\le \theta_\textrm{b} < 112.5\arcdeg$. The bin centers are oriented towards South--East, East, North--East, and North respectively. {\bf While the bin boundaries are arbitrary, they were so chosen to allow bin centers to be aligned along the cardinal and ordinal directions in one half--plane}. Since delays depend on $\theta_\textrm{b}$, binning by $\theta_\textrm{b}$ allows us to match delay spectrum features to different sky directions.

\item {\it Tile Pointing and Power--Pattern}: Since the MWA tiles are steered electronically, the power--pattern of the tile changes in any observing mode that tracks the {\bf pointing center}. In our study, we take into account the effect of the changing tile power--pattern on the delay spectrum (Figure~\ref{fig:power_pattern}). The LST of our data ranges from $\sim$~21 hours through $\sim$~2 hours, and for our analysis we chose the {\it off--zenith} and {\it zenith} pointings shown in Figure~\ref{fig:pointings}. We will show the secondary lobe structures in the tile power--pattern are tied to the foreground signatures in Fourier space.

@@

Figure~\ref{fig:noiseless-dsm-delay-spectrum} shows the amplitude of delay spectrum, $|V_{b\tau}|$, simulated with only the diffuse foreground model. The top and bottom panels correspond to {\it off--zenith} and {\it zenith} pointings, respectively.

\begin{figure}[htb]
\centering
% \includegraphics[width=\linewidth]{figures/v1_0/delta_array_multi_baseline_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_dsm_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
% \includegraphics[width=\linewidth]{f8.eps}
\includegraphics[width=\linewidth]{fig8.eps}
\caption{Simulated delay spectrum {\bf power (in units of K$^2$~(Mpc/$h$)$^3$} for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the diffuse foreground model without any thermal noise. White lines mark the boundaries of the {\it foreground wedge} determined by the horizon delay limit and antenna spacing. The axes and color scale are identical to those in Figures~\ref{fig:fhd_data} and \ref{fig:sim_data}. In the {\it zenith} pointing, delay spectrum from diffuse emission has a {\it two--pronged fork}--shaped structure and is present even at wide antenna spacings. \label{fig:noiseless-dsm-delay-spectrum}}
\end{figure}

@@

The most prominent signature seen in the {\it off--zenith} pointing (top panel of Figure~\ref{fig:noiseless-dsm-delay-spectrum}) is due to the bright Galactic center situated on the western horizon co--located with one of the bright secondary lobes of the power--pattern. It appears as a bright branch in the delay spectrum near the negative delay horizon delay limit. This feature is strongest at short antenna spacings and fades with increasing antenna spacing. The {\bf bright} signature is absent in the {\it zenith} pointing (bottom panel, Figure~\ref{fig:noiseless-dsm-delay-spectrum}) because the Galactic center is below the horizon.

>>>>

Detection of 21~cm emission of neutral hydrogen from the epoch of reionization, at redshifts $z>6$, is limited primarily by foreground emission.  We investigate the chromatic signatures of an all--sky foreground model imprinted by an instrumental transfer function. Using the delay spectrum technique {\bf that maps the Fourier transform of the measurements to locations of foreground objects on the celestial sphere through delays in signals received by antenna pairs}, we demonstrate that the foreground signatures that have the largest impact on the H{\sc i} signal arise from power received far away from the primary field of view. Comparing data from recent Murchison Widefield Array observations with simulations separated into components based on type of emission, we identify diffuse emission near the horizon as a significant contributing factor, even on wide antenna spacings {\bf which are usually not sensitive to diffuse emission}. For signals entering through the primary field of view, compact emission dominates the foreground contamination. These two mechanisms imprint a characteristic {\it pitchfork} signature on the ``foreground wedge'' in Fourier space --- {\bf a signature generic to all wide--field measurements}. Based on these results, we propose that selective down--weighting of data based on antenna spacing vector and time will mitigate foreground contamination substantially (up to a {\bf factor $\sim 100$} with negligible loss of sensitivity) in redshifted 21~cm power spectrum analysis. {\bf A dish--shaped antenna yields the most desirable features in terms of foreground contamination, relative to a dipole or a phased array}.

@@

Considerable effort has been and continues to be made towards understanding the $k$--space behavior of foreground signatures in the observed power spectrum and formulating robust estimators of the true power spectrum \citep{bow09,liu09,dat10,liu11,mor12,tro12,pob13,thy13,dil14,liu14a,liu14b}. {\bf A model that provides a generic explanation for the observed foreground power spectrum has emerged, whereby the wide--field (and chromatic) response of the instrument causes the power in smooth spectrum foregrounds to occupy higher $k$--modes into the so-called `wedge'}. % A general model of scattering of smooth spectrum foregrounds to higher $k$--modes by the wide--field (and chromatic) response of the instrument, into the so-called `wedge', has emerged that provides a high level explanation for the observed foreground power spectrum. 
The conservative foreground strategy that has developed alongside this work is to discard $k$--modes which could be contaminated ({\it avoidance}). The more aggressive alternative is to subtract a sky model and regain access to modes discarded by {\it avoidance}. In both cases, which parts of the sky are most critical to either {\it avoid} or {\it subtract} has remained largely uncertain. Here, we focus primarily on extending the {\it avoidance} strategy by identifying foreground components at greatest risk to ``leak'' from foreground modes to EoR modes and proposing a scheme for down--weighting these components.

@@

After matching observations from the Murchison Widefield Array with instrumental simulations of the entire sky, we examine the {\bf shape of the delay spectrum} of each baseline. We report two important findings: the foregrounds that most severely {\bf obscure} the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather by bright objects from near the horizon; and, diffuse Galactic emission plays a significant role hitherto unpredicted. We quantify these by separating the simulations into components based on type of foreground emission. We then arrive at a new method for minimizing the contribution of bright foregrounds which uses prior knowledge of the sky to down--weight adversely contaminated baselines. {\bf We also examine the role of antenna shapes in determining foreground contamination and provide inputs on antenna design for future EoR experiments}.

In \S\ref{sec:delay-spectrum} we {\bf provide a brief overview of the delay spectrum technique} \citep{par12a,par12b}. In \S\ref{sec:instrument}, we describe the MWA setup, summarize the observing parameters, and present the resulting data. Simulations using these observing parameters are described in \S\ref{sec:modeling} and analyzed by sky location and type of emission in \S\ref{sec:delay-spectrum-analysis}. In \S\ref{sec:fg-grading}, we offer an initial description of a more precise foreground avoidance technique. {\bf The role of antenna shapes in determining foreground contamination is explored in \S\ref{sec:antenna-shape}}. We present a summary of our work and findings in \S\ref{sec:summary}.

@@

Since we are concerned with a redshifted H{\sc i} spectral line from cosmological distances, $f$ is a measure of cosmological distance along the line of sight. $\eta$, which is the Fourier transform dual of $f$, is used to denote the spatial frequency along the line of sight and has units of time. It is directly related to the line--of--sight wavenumber, 
\begin{align}\label{eqn:k-prll}
  k_\parallel &\approx \frac{2\pi\eta\,f_{21}H_0\,E(z)}{c(1+z)^2}, 
\end{align}
where, $f_{21}$ is the rest frame frequency of the 21~cm spin flip transition of H{\sc i}, and $H_0$ and $E(z)\equiv [\Omega_\textrm{M}(1+z)^3+\Omega_\textrm{k}(1+z)^2+\Omega_\Lambda]^{1/2}$ are standard terms in cosmology. This approximation holds under the assumption that the redshift range (or frequency band) is small enough within which cosmological evolution is negligible. Thus,
\begin{align}\label{eqn:los-transform}
  V_{u\eta} &\equiv V_{u\eta}(\boldsymbol{u},\eta) \nonumber\\
  &\equiv \int V_{uf}(\boldsymbol{u},f)\,W'_f\,e^{i2\pi f\eta}\,\dif f
\end{align}
represents the true spatial Fourier representation of the three--dimensional sky brightness distribution. This approach has been discussed in detail in \citet{mor04}. The spatial power spectrum of EoR H{\sc i} distribution, $P(\boldsymbol{k}_\perp,k_\parallel)$, and $V_{u\eta}$ are related by \citep{mor04,mcq06,par12a}: 
\begin{align}\label{eqn:true-power_spectrum}
  P(\boldsymbol{k}_\perp,k_\parallel) &\simeq |V_{u\eta}|^2\left(\frac{A_\textrm{e}}{\lambda^2\Delta B}\right)\left(\frac{D^2\Delta D}{\Delta B}\right)\left(\frac{\lambda^2}{2k_\textrm{B}}\right)^2,
\end{align}
where, $A_\textrm{e}$ is the effective area of the antenna, $\Delta B$ is the bandwidth, $\lambda$ is the wavelength of the band center, and $k_\textrm{B}$ is the Boltzmann constant. Thus, $V_{u\eta}$ inferred from observations, also in units of Jy~Hz, can be converted into an equivalent cosmological H{\sc i} power spectrum $P(\boldsymbol{k}_\perp,k_\parallel)$, in units of K$^2$~Mpc$^3$ or, more generally, K$^2$(Mpc/$h$)$^3$, where $h$ is the Hubble constant factor.

@@

% Radio emission from foreground objects, due to their spatial and spectral (chromatic) structure, contaminate the received signal \citep{bow09,dat10,mor12,tro12,thy13,liu14a,liu14b}. 
Foregrounds can be described in both the delay spectrum and Fourier frameworks. For our study, which primarily concerns with foreground characterization, we find the former to be simple and yet extremely useful, while maintaining a close correspondence with the latter despite subtle differences. In order to express a quantity derived from $V_{u\eta}$ that has units identical to that of aforementioned cosmological H{\sc i} power spectrum, we define:
\begin{align}\label{eqn:fake-power-spectrum}
  P_\textrm{d}(\boldsymbol{k}_\perp,k_\parallel) &\equiv |V_{b\tau}|^2\left(\frac{A_\textrm{e}}{\lambda^2\Delta B}\right)\left(\frac{D^2\Delta D}{\Delta B}\right)\left(\frac{\lambda^2}{2k_\textrm{B}}\right)^2,
\end{align}
where, $\boldsymbol{u}$ and $\eta$ in equations~\ref{eqn:k-perp} and \ref{eqn:k-prll} have been correspondingly replaced with $\boldsymbol{b}/\lambda$ and $\tau$ respectively.

In summary, the delay spectrum, $V_{b\tau}$, is obtained from {\it visibilities} which are the basic data blocks measured by each antenna pair, using equations~\ref{eqn:obsvis} and \ref{eqn:delay-transform}. $V_{b\tau}$ captures all the effects of EoR H{\sc i} signal corruption caused by foregrounds and the instrument. At the same time, it is closely related to the sought power spectrum $P(\boldsymbol{k}_\perp,k_\parallel)$ containing critical information about spatial scales. In this paper, we often use the quantity $P_\textrm{d}(\boldsymbol{k}_\perp,k_\parallel)$, and sometimes $V_{b\tau}$.

@@

For geometrical intuition, we restrict the orientation ($\theta_b$, measured anti-clockwise from East) of all baselines to lie in the range $-67.5\arcdeg \leq \theta_b < 112.5\arcdeg$. Baselines oriented in the other half--plane measure conjugate visibilities with delays of equal magnitude but of opposite sign and hence are ignored in our analysis. {\bf Note that visibilities from different baselines have not been averaged together in this analysis}.

@@

The {\it local sidereal time} (LST) at these two snapshots are shown in Figure~\ref{fig:pointings} as vertical lines at {\bf 22.08~hours (wrapped to $-1.92$~hours in the figure)} and 0.09~hours, which are hereafter denoted as {\it off--zenith} and {\it zenith} pointings, respectively.

@@

The data are flagged\footnote{\url{http://sourceforge.net/p/aoflagger}} for interference \citep{off10,off12}, removing 3\% of the data and averaged in time  and frequency from the raw 0.5~s, 40~kHz to 2s, 80~kHz. These data are then calibrated to a simulation of the sky containing 2420 point--like objects selected from the MWA Commissioning Survey \citep[MWACS;][]{hur14}. It has a flux density limit of 25~mJy and a declination range of $-12\arcdeg$~ to $-40\arcdeg$~ evenly covering the field of view of the observations reported here. The objects used in calibration are selected to lie inside the 5\% contour of the primary lobe of the tile power--pattern. The calibration algorithm -- based on forward modeling software by \citet{sul12} and the calibration method described by \citet{sal14} -- computes complex gain solutions per channel per antenna averaged to two minute intervals. The solutions are fairly low signal--to--noise so we iteratively average along the antenna and frequency dimensions to capture the relatively independent passband and antenna--to--antenna variation. First, we average the channel gains over all antennas to obtain a high signal--to--noise measurement of the bandpass. After applying this single passband, we do a second round of calibration and fit second and first order polynomials for amplitude and phase respectively for each antenna. This flattens any residual variation in bandpass and removes small phase slopes due to variations in cable delay. Finally, we fit for an additional phase known to be caused by small reflections in a subset of cables.

@@

% In subsequent sections, we plot the amplitude of the delay spectrum, $|V_{b\tau}|$, sorted by baseline length, $|\boldsymbol{b}|$. Note that visibilities from different baselines have not been averaged together in this analysis.

We show {\bf delay power spectra} from MWA data after deconvolution along the delay axis in Figure~\ref{fig:fhd_data}. The principle feature is the {\it foreground wedge}, whose boundaries, marked by white lines, increase linearly with light travel time corresponding to baseline length. Note that the wedge is not filled out uniformly. Each baseline has a different orientation and resulting response to different objects in the sky. It is this variation that we will explore below.

@@

\begin{figure}[htb]
\centering
% \includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_fhd_delay_spectrum_snapshots.eps}
% \includegraphics[width=\linewidth]{f3.eps}
\includegraphics[width=\linewidth]{fig3.eps}
\caption{{\bf Delay power spectra} of two observations of the EoR field (RA = 0$^\textrm{h}$, Dec = $-30\arcdeg$) at an {\it off--zenith} pointing (top) when tiles have significant secondary lobes in the direction of the Galactic center and a {\it zenith} pointing (bottom) where secondary lobes are minimal and the Galaxy has set. {\bf In both cases, the delay spectra are phased to the zenith and thus the {\it foreground wedge} boundaries are symmetric around $\tau=0$}. For reference, the $k$--space equivalent of $\tau$ and $|\boldsymbol{b}|$, namely, $k_\parallel$ and $k_\perp$ are shown at the axes on the right and top respectively. White lines mark the boundaries of {\it foreground wedge} determined by light travel time across each baseline. Smooth spectrum objects predominantly reside in this {\it foreground wedge}. The logarithmic color scale (shown at the right) is common to both panels. \label{fig:fhd_data}}
\end{figure}

@@

{\bf Estimation of $T_\textrm{sys}$ from MWA data is a subject of active investigation. For our work, we estimate thermal noise in the data using the {\it rms} of $V_{b\tau}$ obtained after {\it delay--deconvolution} across all antenna spacings for $|\tau| \geq 1$~$\mu$s using the relations:
\begin{align}\label{eqn:Tsys}
  \Delta V_{b\tau}^\textrm{rms} &= \sqrt{N_\textrm{ch}}\,\Delta V_{bf}^\textrm{rms} \Delta f, \textrm{and} \\
  \Delta V_{bf}^\textrm{rms} &= \frac{2\,k_\textrm{B}\,T_\textrm{sys}}{A_\textrm{e}\sqrt{2\,\Delta f\,\Delta t}},
\end{align}
where, $\Delta f = 80$~kHz, $\Delta t = 112$~s, and $N_\textrm{ch}=\Delta B/\Delta f$ is the number of frequency channels. The choice of threshold for $\tau$ is well outside the {\it foreground window}, where foreground contamination is negligible and thus yields a robust estimate of $T_\textrm{sys}$. We find the average system temperature to be $\sim 95$~K across all frequency channels and across all antenna pairs throughout the course of the observation. Hence, for our simulations, we use $T_\textrm{sys}=95$~K to match the thermal noise observed in the data.}

@@

With the aforementioned foreground model, and instrumental and observational parameters, we simulate visibilities using equation~\ref{eqn:obsvis}. Figure~\ref{fig:sim_data} shows the {\bf delay power spectra} from {\it off--zenith} and {\it zenith} pointings. Notice the qualitative agreement of amplitude and structure with those obtained from data shown in Figure~\ref{fig:fhd_data}. The Galactic center and the Galactic plane visible in the {\it off--zenith} pointing make it appear brighter in the {\it foreground wedge}.

@@

In order to make a quantitative comparison of delay spectra obtained with MWA data and our simulations, we consider the uncertainty in the assumed spectral index of our foreground model. Our foreground models are derived from other higher frequency catalogs and sky maps. The inherent spread in spectral index increases the uncertainty while predicting fluxes at the observing frequency. Using simple error propagation, the fractional error in the delay spectrum {\bf amplitude}, $|V_{b\tau}|$, caused by the spread in spectral index is $\sim \ln(f_\textrm{orig}/f)\,\Delta\alpha_\textrm{sp}$, where, $f_\textrm{orig}$ is the original frequency at which the catalog or map was created, $f=185$~MHz is the MWA observing frequency, and $\Delta\alpha_\textrm{sp}$ is the spread (HWHM) in spectral index. From \citet{mau03}, we assume $\Delta\alpha_\textrm{sp} \approx 0.35$ for point sources from NVSS and SUMSS catalogs. Although the model of \citet{deo08} yields a spectral index per direction on the sky, we could assume similar uncertainties exist in spectral indices of our diffuse sky model as well, which is predominantly derived from the 408~MHz map of \citet{has82}. Thus, fractional errors in delay spectrum {\bf amplitudes} from compact and diffuse components are $\sim$70\% and $\sim$30\% respectively.

@@

\item {\it Baseline Orientation}: Since the spatial structure of our foreground model is not expected to be isotropic, we divide our antenna spacings by their orientation, $\theta_\textrm{b}$. We use the {\bf four following bins sweeping 45\arcdeg each}: $-67.5\arcdeg\le \theta_\textrm{b} < -22.5\arcdeg$, $-22.5\arcdeg\le \theta_\textrm{b} < 22.5\arcdeg$, $22.5\arcdeg\le \theta_\textrm{b} < 67.5\arcdeg$, and $67.5\arcdeg\le \theta_\textrm{b} < 112.5\arcdeg$. The bin centers are oriented towards South--East, East, North--East, and North respectively. {\bf While the bin boundaries are arbitrary, they were so chosen to allow bin centers to be aligned along the cardinal and ordinal directions in one half--plane}. Since delays depend on $\theta_\textrm{b}$, binning by $\theta_\textrm{b}$ allows us to match {\bf features of the delay power spectra} to different sky directions.

\item {\it Tile Pointing and Power--Pattern}: Since the MWA tiles are steered electronically, the power--pattern of the tile changes in any observing mode that tracks the {\bf pointing center}. In our study, we take into account the effect of the changing tile power--pattern on the {\bf delay power spectra} (Figure~\ref{fig:power_pattern}). The LST of our data ranges from $\sim$~21 hours through $\sim$~2 hours, and for our analysis we chose the {\it off--zenith} and {\it zenith} pointings shown in Figure~\ref{fig:pointings}. We will show the secondary lobe structures in the tile power--pattern are tied to the foreground signatures in Fourier space.

@@

Figure~\ref{fig:noiseless-dsm-delay-spectrum} shows the {\bf delay power spectra} simulated with only the diffuse foreground model. The top and bottom panels correspond to {\it off--zenith} and {\it zenith} pointings, respectively.

\begin{figure}[htb]
\centering
% \includegraphics[width=\linewidth]{figures/v1_0/delta_array_multi_baseline_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_dsm_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
% \includegraphics[width=\linewidth]{f8.eps}
\includegraphics[width=\linewidth]{fig8.eps}
\caption{Simulated {\bf delay power spectra (in units of K$^2$~(Mpc/$h$)$^3$} for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the diffuse foreground model without any thermal noise. White lines mark the boundaries of the {\it foreground wedge} determined by the horizon delay limit and antenna spacing. The axes and color scale are identical to those in Figures~\ref{fig:fhd_data} and \ref{fig:sim_data}. In the {\it zenith} pointing, {\bf delay power spectrum} from diffuse emission has a {\it two--pronged fork}--shaped structure and is present even at wide antenna spacings. \label{fig:noiseless-dsm-delay-spectrum}}
\end{figure}

@@

The most prominent signature seen in the {\it off--zenith} pointing (top panel of Figure~\ref{fig:noiseless-dsm-delay-spectrum}) is due to the bright Galactic center situated on the western horizon co--located with one of the bright secondary lobes of the power--pattern. It appears as a bright branch in the {\bf delay power spectrum} near the negative delay horizon delay limit. This feature is strongest at short antenna spacings and fades with increasing antenna spacing. The {\bf bright} signature is absent in the {\it zenith} pointing (bottom panel, Figure~\ref{fig:noiseless-dsm-delay-spectrum}) because the Galactic center is below the horizon.
